  ---
title: "Querying iAtlas Data"
output: html_notebook
---

The data hosted by iAtlas is available through an API.

```{r}
# We have a few libraries to install.
try({
  packages = c("magrittr", "dplyr", "tidyr", "dplyr", "tidyr", "ggplot2")

  sapply(packages, function(x) {
    if (!require(x,character.only = TRUE))
      install.packages(x)
    suppressPackageStartupMessages(library(x,character.only = TRUE))
  })},
  silent=TRUE
)

# and our iatlas package from github
if (!require(iatlasGraphQLClient)) {
  devtools::install_github("CRI-iAtlas/iatlasGraphQLClient")
  library(iatlas.modules)
}

# helper functions
# git clone the notebook repo to get this file #
# or see: https://github.com/CRI-iAtlas/iatlas-notebooks/blob/main/functions/notebook_functions.R
# setwd('C:/Users/12064/DataspellProjects/iatlas-notebooks')
source('functions/notebook_functions.R')
```

Exploring the datasets and features

The iAtlas data is stored in a database that can be queried with functions from the
iatlasGraphQLClient package. We have clinical data, immune features, scores of
predictors of response to immunotherapy, and quantile normalized gene expression.
You can get more information in iAtlas on immune features, and our annotation of
immunomodulators genes. You can access more information about these
datasets in iAtlas.

As a first step, let’s take a look at the available datasets.

```{r}
# datasets that we have in the iAtlas database
datasets <- iatlasGraphQLClient::query_datasets()
datasets
```

Let's count the datasets by type:

```{r}
table(datasets$type)
```

Data is also tagged with different descriptive terms which can help
in finding appropriate data given a particular interest.

```{r}
# datasets that we have in the iAtlas database
dataset_tags <- iatlasGraphQLClient::query_dataset_tags()
table(dataset_tags$tag_name)
```

Suppose, we're interested in response data.

```{r}
dataset_tags[dataset_tags$tag_name == 'TCGA_Study',]
```

From that list, we might be interested in a particular dataset.
In order to get the sample IDs associated with that dataset, we use
the query_dataset_samples function.

```{r}
dataset_samples <- iatlasGraphQLClient::query_dataset_samples(datasets = "TCGA")
dim(dataset_samples)
```

OK, so, there's 11,080 samples in the TCGA dataset. Let's see the top of the dataset_samples table.

```{r}
head(dataset_samples)
```


```{r}
d = iatlasGraphQLClient::query_cohorts()
d[d$dataset_name == 'TCGA',]
```
```{r}
tcga_studies <- iatlasGraphQLClient::query_tags_with_parent_tags(parent_tags = "TCGA_Study")
tcga_studies
```
```{r}
d = iatlasGraphQLClient::query_tag_samples(parent_tags = "TCGA_Study",tags= c(tag_name = "KICH"))
d
```
