---
title: "PRINCE Study Data Plots"
output: html_notebook
---

Install needed packages
```{r}
try({
    packages = c("magrittr", "dplyr", "feather", "tidyr", "survival", "survminer", "plotly", "repr", "purr")
    sapply(packages, function(x) {
      if (!require(x,character.only = TRUE))
        install.packages(x)
        library(x,character.only = TRUE)
    })

    if (!require(iatlas.modules)) {
        devtools::install_github("th86/concordanceIndex")
        library(concordanceIndex)
    }
    
    # and our iatlas package from github
    if (!require(iatlas.modules)) {
      devtools::install_github("CRI-iAtlas/iatlas.modules")
      library(iatlas.modules)
    }},
    silent = TRUE
)

```



Source needed functions. These are functions from the iAtlas app source code, copied here for convenience
```{r}
source('functions/notebook_functions_2.R')
```

There are other notebooks in this repository that illustrate functions used to query the iAtlas database to obtain data. We use these functions below.

### Build Data Matrices
Build data frame of feature values. Includes e.g. gene expression signature scores and survival
```{r}
## use this if you just want to see available features
##features_df <- iatlasGraphQLClient::query_features(cohorts = "PRINCE") 
feature_value_df <- iatlasGraphQLClient::query_feature_values(cohorts = "PRINCE")
feature_value_df <- feature_value_df %>% rename(sample_name=sample) ## expected by later functions
feature_value_df <- feature_value_df %>% mutate(dataset_name="PRINCE") ## expected by later functions
```

Obtain samples IDs for the PRINCE study
```{r}
prince_samples  <- iatlasGraphQLClient::query_dataset_samples(datasets = "PRINCE")$sample_name
```

Build sample matrix containing patient, ICI treatment, and response
```{r}
m1 <- iatlasGraphQLClient::query_sample_patients(samples=prince_samples) %>% select(sample_name,patient_name)
tags <- c("ICI_Rx","Responder","Response","Sample_Treatment")
m3 <- iatlasGraphQLClient::query_tag_samples_parents(cohort="PRINCE",parent_tags = tags) %>% distinct()
m4 <- m3 %>% select(sample_name,parent_tag_name,tag_short_display)
m5 <- m4 %>% pivot_wider(names_from = parent_tag_name,values_from= tag_short_display)
sample_mat <- m5 %>% inner_join(m1,by="sample_name")
```

Counts
```{r}
cat("No of samples:",nrow(sample_mat),"\n")
cat("Responder:")
table(sample_mat$Responder)
cat("Response:")
table(sample_mat$Response)
cat("ICI_Rx:")
table(sample_mat$ICI_Rx)
cat("Sample_Treatment:")
table(sample_mat$Sample_Treatment)
```
We'll build some relevant sample subsets here. In the iAtlas app, these can be made using the Cohort selection. 

Sample IDs for samples obtained prior to ICI treatment, and during treatment
```{r}
samps_pre <- sample_mat %>% filter(Sample_Treatment=="Pre-treatment") %>% purrr::pluck("sample_name")
samps_during <- sample_mat %>% filter(Sample_Treatment=="During treatment") %>% purrr::pluck("sample_name")
```

Sample IDs for samples where the response is known
```{r}
samps_known_response <- sample_mat %>% filter(Response != "Not available") %>% purrr::pluck("sample_name")
```

##
## Survival plot
##
We construct a data frame using the iAtlas app function `build_survival_df` and plot the survival curves using another function `create_kmplot`.

The primary sample division is by Responder, so let's make a data frame of those values 
```{r}
df_responder <- iatlasGraphQLClient::query_tag_samples_parents(cohort="PRINCE",parent_tags = "Responder") %>% distinct() %>%
select(sample_name,tag_short_display) %>% rename(group_name=tag_short_display)
```

Join that to the data frame of feature values (which has the survival data)
```{r}
df <- feature_value_df %>% left_join(df_responder,by="sample_name")
```

The secondary division of the samples will be by the Macrophage Regulation signature. Create data frame with that.
```{r}
extra_group_df <- feature_value_df %>% filter(feature_name=="CSF1_response") 
```

Restrict to pre-treatment samples and corresponding patients for which response is known
```{r}
df <- df %>% filter(sample_name %in% intersect(samps_pre,samps_known_response))
extra_group_df <- extra_group_df %>% filter(sample_name %in% intersect(samps_pre,samps_known_response))
```

Build the data frame going that goes into the plot. Here there is an option to include the extra sample group or not. 
```{r}
df_new <- build_survival_df(df,group_column="group_name", time_column="PFI_time_1", filter_df = TRUE,  extra_group_df , extra_group = "No")
df_surv <- df_new %>% distinct()
```

We end up with time, status, measure. In iatlas-app this is done with a shiny reactive
```{r}
fit <- survival::survfit(survival::Surv(time, status) ~ measure, data = df_surv)
```


```{r,fig.width=10, fig.height=6}

create_kmplot(fit, 
              df_surv, 
              confint=FALSE, 
              risktable=TRUE, 
              title="PRINCE - PDAC, PD-1", 
              group_colors=NULL, 
              show_pval = TRUE, 
              show_pval_method = FALSE, 
              facet = FALSE)
```

### Immunomodulators

```{r}
immunomodulators <- iatlasGraphQLClient::query_immunomodulators()
head(immunomodulators)
```

Immunomodulator values 
```{r}
gene_vals <- iatlasGraphQLClient::query_gene_expression(cohorts = "PRINCE", entrez = immunomodulators$entrez)
```

```{r}
im_exp <- gene_vals %>% select(-entrez) %>% pivot_wider(names_from = hgnc,values_from=rna_seq_expr, values_fill=NA) %>% rename(sample_name=sample)
```

```{r}
m <- sample_mat %>% inner_join(im_exp,by="sample_name")
```

```{r}
mm <- m %>% group_by(Responder) %>% summarise(across(c(`HLA-DRA`:`HLA-A`),mean)) %>% as.data.frame()
g <- mm[,2:ncol(mm)] ; rownames(g) <- mm[,1]
g <- t(g)
```

```{r}
g %>% ggplot(aes(`Non-Responder`,Responder,label=rownames(.))) + geom_point() + scale_x_log10()  + scale_y_log10() + geom_abline()+  geom_text()  
```


```{r}
mm <- m %>% group_by(`ICI_Rx`) %>% summarise(across(c(`HLA-DRA`:`HLA-A`),mean)) %>% as.data.frame()
g <- mm[,2:ncol(mm)] ; rownames(g) <- mm[,1]
g <- t(g)
```




```{r}
g %>% ggplot(aes(`Nivolumab`,`None`,label=rownames(.))) + geom_point() + scale_x_log10()  + scale_y_log10() + geom_abline()+  geom_text()  
```






